---
title: "Assessment-HarvadX-Combinations-and-Permutations"
output: html_notebook
---

# NOTEBOOK BY ISABELLA HEDER

# **Assessment HarvardX: Combinations and Permutations - Data Science: Probability course**

## **EXERCISES:**

## Exercise 1:

Two teams, say the Celtics and the Cavs, are playing a seven game series. The Cavs have a 60% chance of winning each game. What is the probability that the Celtics win at least one game? Remember that the Celtics must win one of the first four games, or the series will be over!

1.  **Instructions:**

    -   Calculate the probability that the Cavs will win the first four games of the series.

    -   Calculate the probability that the Celtics win at least one game in the first four games of the series.

```{r}
# Use `p_cavs_win4` as the probability that the Cavs will win the first four games of the series.
p_cavs_win4 <-  0.6 * 0.6 * 0.6 * 0.6 # or 0.6^4

# Using the variable `p_cavs_win4`, calculate the probability that 
# the Celtics win at least one game in the first four games of the series.
1 - p_cavs_win4
```

The solution was based in the **Multiplication Rule** for **independent** events

**P(A, B and C) = P(A).P(B).P(C)**

The Cavs had 60% chance of winning EACH game, no matter the outcome of the other games.

> For the second part:

**If the Cavs lose one game = the Celtics win one.**

That's why I used:

1 (meaning 100%) - probability that the cavs will win the first four = probability that the celtics will win at least one game

## Exercise 2:

Create a Monte Carlo simulation to confirm your answer to the previous problem by estimating how frequently the Celtics win at least 1 of 4 games. Use B ← 10000 simulations. The provided sample code simulates a single series of four random games, simulated_games.

1.  **Instructions:**

    -   Use the replicate function for B ← 10000 simulations of a four game series. The results of replicate should be stored in a variable named celtic_wins.

    -   Within each simulation, replicate the sample code to simulate a four-game series named simulated_games. Then, use the any function to indicate whether the four-game series contains at least one win for the Celtics. Perform these operations in two separate steps.

    -   Use the mean function on celtic_wins to find the proportion of simulations that contain at least one win for the Celtics out of four games.

```{r}
# Let's run the Monte Carlo simulation 10,000 times.
B <- 10000

# Use the `set.seed` function to make sure your answer matches 
# the expected result after random sampling.
set.seed(1)

# Create an object called `celtic_wins` that replicates two steps 
# for B iterations: (1) generating a random four-game 
# series named `simulated_games`, then (2) determining 
# whether the simulated series contains at least one win for the Celtics.

# Calculate the frequency out of B iterations that the Celtics won at least one game.
celtic_wins <- replicate(B, {
  simulated_games <- sample(c("lose", "win"), 4, replace = TRUE, prob = c(0.6, 0.4))
  any(simulated_games == "win")})

mean(celtic_wins)

```

The solution was based in **Monte Carlo Simulations** for Celtics winning a game, using the *same values as the previous exercise*.

B = number of times the exercise wanted the Monte Carlo simulation to run

**Celtic_wins:**

Using **replicate** function to repeat the simulation for the number of games.

Using **sample** function to randomly select outcomes (win or lose) for each game.

-   4 games, each with 40% chance of winning and 60% chance of losing.

**Replace = TRUE;** means that every time the simulation selects a value, it goes back to the mass, not taking it away from the total.

Using **any( )** to check if at least one game received the "win" value.

--

Using **mean( )** to calculate the proportion of simulations where the Celtics won at least one game.

--

## Exercise 3:

Say you’ve drawn 5 balls from a box that has 3 cyan balls, 5 magenta balls, and 7 yellow balls, with replacement, and all have been yellow. What is the probability that the next one is yellow?

1.  **Instructions:**

    -   Assign the variable p_yellow as the probability of choosing a yellow ball on the first draw.

    -   Using the variable p_yellow, calculate the probability of choosing a yellow ball on the sixth draw.

```{r}
# Assign the variable 'p_yellow' as the probability that a 
# yellow ball is drawn from the box.
box <- rep(c('cyan', 'magenta', 'yellow'), times = c(3, 5, 7))
p_yellow <- mean(box == 'yellow')
print(p_yellow)

# Using the variable 'p_yellow', calculate the probability of 
# drawing a yellow ball on the sixth draw. Print this value to the console.
print(p_yellow)

```

For this exercise I decided to **create a vector** using the amount of balls by color.

To calculate the probability of choosing a yellow ball I used the mean( ) function, to calculate the proportion of yellow balls over the total.

It would be the same as: 7/15

To calculate the probability of drawing a yellow ball on the sixth draw, knowing that they’ve drawn 5 balls from the box and all have been yellow, I used the **same proportion!**

Because this exercise was done **WITH replacement**, meaning, every time a ball was drawn, it would immediately go back to the box.

## Exercise 4:

If you roll a 6-sided die once, what is the probability of not seeing a 6? If you roll a 6-sided die six times, what is the probability of not seeing a 6 on any of those rolls?

1.  **Instructions:**

    -   Assign the variable p_no6 as the probability of not seeing a 6 on a single roll.

    -   Then, calculate the probability of not seeing a 6 on six rolls using p_no6.

```{r}
# Assign the variable 'p_no6' as the probability of not seeing 
# a 6 on a single roll.
p_no6 <- 5/6
p_no6

# Calculate the probability of not seeing a 6 on 
# six rolls using `p_no6`. Print your result to the console: 
# do not assign it to a variable.
p_no6^6
```

For the final exercise:

I was asked to calculate the probability of not seeing a 6 after rolling a 6-sided die one time only.

Since the die has 6 sides: only one of them has the value "6"

> So not seeing a 6 means I would see any of the other numbers (1, 2, 3, 4 and 5) –\> that would mean 5/6

**Not seeing a 6 on 6 rolls:**

I used the **multiplication rule** for **independent** events, because each roll does not change because of the other outcomes.

**P(A, B, C, D, E and F) = P(A) . P(B) . P(C) . P(D) . P(E) . P(F)**
